---
title: "PM 566 Lab 3"
author: "Erica Shin"
format: html
editor: visual
embed-resources: true
---

## Exercise 1

```{r}
#install.packages('R.utils')

library(R.utils)

download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
met <- as.data.frame(met)
```

## Exercise 2

```{r}
dim(met)
nrow(met)
ncol(met)
```

There are 2,377,343 rows and 30 columns.

## Exercise 3

```{r}
str(met)
```

Variables of interest are: year, day, hour, elev, temp, and wind.sp.

## Exercise 4

```{r}
table(met$year)

table(met$day)

table(met$hour)

summary(met$temp)

summary(met$elev)

summary(met$wind.sp)

met[met$elev==9999.0, "elev"] <- NA
summary(met$elev)

met <- met[met$temp > -40, ]
head(met[order(met$temp), ])
```

There are 710 missing values in the wind speed variable.

## Exercise 5

The suspicious temperature value of -17.2C has a location with a latitude of 38.767 and a longitude of -104.3. This is located in Yoder, Colorado.

It does not seem reasonable that Yoder, Colorado would have a temperature reading of -17.2C in August.

This location is near the USAF Academy Bullseye Auxiliary Airfield in Yoder, Colorado, which has an elevation of approximately **6,036 ft (1,840 m)**. Therefore, the range of elevations make sense (-13m to 4113m).

## Exercise 6

```{r}
elev <- met[which(met$elev == max(met$elev, na.rm = TRUE)), ]
summary(elev)

cor(elev$temp, elev$wind.sp, use="complete")

cor(elev$temp, elev$hour, use="complete")

cor(elev$wind.sp, elev$day, use="complete")

cor(elev$wind.sp, elev$hour, use="complete")

cor(elev$temp, elev$day, use="complete")
```

## Exercise 7

```{r}
#install.packages("leaflet")

library(leaflet)

leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)

library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

elev <- elev[order(elev$date), ]
head(elev)

#histograms of the elevation, temperature, and wind speed variables for the whole dataset


hist(met$elev)

hist(met$temp)

hist(met$wind.sp)

#line graphs of temperature vs. date and wind speed vs. date

plot(met$temp, met$date)

plot(met$wind.sp, met$date)
```

## Exercise 8

What does "qc" mean for some of the data variable names?
